---
title: "powerlaw"
author: "jmm"
date: "September 20, 2015"
output: html_document
---

Some notes on plotting power law data moslty taken from [Newman 2006](http://arxiv.org/pdf/cond-mat/0412004.pdf). 

We can simulate some data from a power law distribution using an inverse appoach:
```{r}
rpow <- function(n, mu = -2, ymin = 1, ymax = 10e4){
  x <- runif(n, 0, 1)
  y <- ( (ymax^(mu+1) - ymin^(mu+1)) * x + ymin^(mu+1) )^(1/(mu+1))
}

ymin <- 60
ymax <- 10e4
y <- rpow(10000, mu = -1.5, ymin = ymin, ymax = ymax)
```
One option is to plot a histogram on log log scale but this results in noisy pattern at large scales:

```{r}
tmp <- hist(y, 200)
plot(tmp$mids, tmp$counts, log = "xy")
```

A good alternative is to use a log scaled binning in the histogram (and then normalizing counts by bin size):

```{r}
require(emdbook)
bins <- lseq(ymin, ymax, length.out = 100)
tmp <- hist(y, breaks = bins)
binsize <- bins[2:length(bins)] - bins[1:(length(bins) - 1)]
normcounts <- tmp$counts / binsize
plot(tmp$mids, normcounts, log = "xy")
```

By far the best alternative is to plot the "rank frequency distribution"

```{r}
  
ys <- sort(y)
n <- length(y)
vals <- unique(ys)
cs <- cumsum(tabulate(match(ys, vals)))/n
  
#plot(vals, 1-cs, log = "xy", type = "l")

plot(vals, 1-cs, type="l", log="xy", axes = FALSE)
axis(side = 2, at = (locs <- 1/c(1,10,100,1000, 10000, 100000)), labels = locs, las=1)
axis(side = 1, at = (locs <- c(1,10,100,1000, 10000, 100000)))

```